---
title: "p8105_hw1_jy2965"
author: "Jun Yin"
date: 2019-09-19
output: github_document
---
## Problem 1

## Create a data frame
The chunk below creates a dataframe containing a random sample of size 8 from a standard Normal distribution, a logical vector indicating whether elements of the sample are greater than 0, a character vector of length 8 and a factor vector of length 8, with 3 different factor "levels".


```{r homework_problem_1_dataframe}
library(tidyverse)
set.seed(1)
problem1_df = tibble(
  vec_numeric = rnorm(8),
  vec_logical = vec_numeric>0,
  vec_char = c("My", "name", "is", "Jun","trying","to","get","eight"),
  vec_factor = factor(c("chocolate", "chocolate", "cookie", "cookie","cookie","cake","cake","cake"))
)
class(problem1_df$vec_numeric)
class(problem1_df$vec_logical)
class(problem1_df$vec_char)
class(problem1_df$vec_factor)
levels(problem1_df$vec_factor)


mean(problem1_df$vec_numeric)
mean(problem1_df$vec_logical)
mean(problem1_df$vec_char)
mean(problem1_df$vec_factor)

```
Result: I could get the mean of numeric and logical variables, but I can not get the mean of character and factor variables.

Explanation: Because the numeric variable is a set of numbers, it is easy to calculate the mean of it. As for logical variables, the system recognizes TRUE as 1 and FALSE as 0, so it can calculate the mean of a set of number 1 and 0.
However, sice character and factor variables are sets of words instead of numbers, the system cannot calculate the means of them.


## data conversion

This chunk shows the conversion of different types of variables.
```{r homework_problem_1_dataconvert, warning = FALSE, results = FALSE}
as.numeric(problem1_df$vec_logical)#convert the logical vector to numeric
as.numeric(problem1_df$vec_logical)*problem1_df$vec_numeric#multiply the random sample by results
as.factor(problem1_df$vec_logical)#convert the logical vector to a factor
as.factor(problem1_df$vec_logical)*problem1_df$vec_numeric#multiply the random sample by results
as.numeric(as.factor(problem1_df$vec_logical))#convert the logical vector to a factor and then convert the result to numeric
as.numeric(as.factor(problem1_df$vec_logical))*problem1_df$vec_numeric#multiply the random sample by results
```
Results: I could get a valid results when I converted the logical variable to numeric variable and multiplied random sample by results. Also I could get a valid results when I converted the logical variable to factor variable and then convert the result to numeric variable before I multiplied random sample by results. However, I can not get a valid results when I multiplied the randome sample by the results of converting logical variable to factor variable.
Because after we convert other types of variable to numeric variable, there will be a set of number which could be calculated by multiplying a random sample. And factor variables is a set of words instead of numbers, the system cannot calculate them by multiplying a random sample.
I think it does help explain what happens when I try to take the mean.


## Problem 2

# create datafram
This data fram comprised of:
x: a random sample of size 500 from a standard Normal distribution
y: a random sample of size 500 from a standard Normal distribution
A logical vector indicating whether x + y > 1
A numeric vector created by coercing the above logical vector
A factor vector created by coercing the above logical vecto

```{r homework_problem_2_dataframe}
problem2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec2_logical = x+y > 1,
  vec2_numeric = as.numeric(vec2_logical),
  vec2_factor = as.factor(vec2_logical) 
)
```
## inline R code
As the data frame shows above, *the number of the rows of the dataset* is `r nrow(problem2_df)`, *the number of the column of the dataset* is `r ncol(problem2_df)`; *the mean of x* is `r mean(problem2_df$x)`, *the median of x* is`r median(problem2_df$x)`, *the standard divation of x* is `r sd(problem2_df$x)`; Besides, *the proportion of cases for which x + y > 1* is `r sum(problem2_df$vec2_logical)/length(problem2_df$vec2_logical)`.

##scatterplot
### scatterplot of y vs x with the color points using the logical variable
```{r Scatterplot_logical}
ggplot(problem2_df, aes(x = x, y = y, color = problem2_df$vec2_logical)) + geom_point()
ggsave("scatterplot_with_color_points_of_logical_variable.png")
```
From the scatterplot above, we could find that nearly 50% of the logical variables are true and 50% are false.

### scatterplot of y vs x with the color points using the numeric variable
```{r Scatterplot_numeric}
ggplot(problem2_df, aes(x = x, y = y, color = problem2_df$vec2_numeric)) + geom_point()
```
From the scatterplot above, we could find that nearly 50% of the numeric variables equal 1 and 50% equal 0.

### scatterplot of y vs x with the color points using the factor variable
```{r Scatterplot_factor}
ggplot(problem2_df, aes(x = x, y = y, color = problem2_df$vec2_factor)) + geom_point()
```
From the scatterplot above, we could find that nearly 50% of the factor variables are ture and 50% are false.
