---
title: "p8105_hw1_jy2965"
author: "Jun Yin"
date: 2019-09-19
output: html_document
---
## Problem 1

## Create a data frame
The chunk below creates a dataframe containing a random sample of size 8 from a standard Normal distribution, a logical vector indicating whether elements of the sample are greater than 0, a character vector of length 8 and a factor vector of length 8, with 3 different factor "levels".


```{r homework_problem_1_dataframe}
library(tidyverse)
set.seed(1)
problem1_df = tibble(
  vec_numeric = rnorm(8),
  vec_logical = vec_numeric>0,
  vec_char = c("My", "name", "is", "Jun","trying","to","get","eight"),
  vec_factor = factor(c("chocolate", "chocolate", "cookie", "cookie","cookie","cake","cake","cake"))
)
class(problem1_df$vec_numeric)
class(problem1_df$vec_logical)
class(problem1_df$vec_char)
class(problem1_df$vec_factor)
levels(problem1_df$vec_factor)


mean(problem1_df$vec_numeric)
mean(problem1_df$vec_logical)
mean(problem1_df$vec_char)
mean(problem1_df$vec_factor)

```
Result: I could get the mean of numeric and logical variables, but I can not get the mean of character and factor variables.

Explanation: Because the numeric variable is a set of numbers, it is easy to calculate the mean of it. As for logical variables, the system recognizes TRUE as 1 and FALSE as 0, so it can calculate the mean of a set of number 1 and 0.
However, sice character and factor variables are sets of words instead of numbers, the system cannot calculate the means of them.


## data conversion

This chunk shows the conversion of different types of variables.
```{r homework_problem_1_dataconvert, warning = FALSE, results = FALSE}
as.numeric(problem1_df$vec_logical)#convert the logical vector to numeric
as.numeric(problem1_df$vec_logical)*problem1_df$vec_numeric#multiply the random sample by results
as.factor(problem1_df$vec_logical)#convert the logical vector to a factor
as.factor(problem1_df$vec_logical)*problem1_df$vec_numeric#multiply the random sample by results
as.numeric(as.factor(problem1_df$vec_logical))#convert the logical vector to a factor and then convert the result to numeric
as.numeric(as.factor(problem1_df$vec_logical))*problem1_df$vec_numeric#multiply the random sample by results
```
Results: I could get a valid results when I converted the logical variable to numeric variable and multiplied random sample by results. Also I could get a valid results when I converted the logical variable to factor variable and then convert the result to numeric variable before I multiplied random sample by results. However, I can not get a valid results when I multiplied the randome sample by the results of converting logical variable to factor variable.
Because after we convert other types of variable to numeric variable, there will be a set of number which could be calculated by multiplying a random sample. And factor variables is a set of words instead of numbers, the system cannot calculate them by multiplying a random sample.
I think it does help explain what happens when I try to take the mean.
```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
